---
title: "Visualization #2"
description: |
  Strength of memory representations depend on learning goals.
author:
  - name: Lea Frank
    url: https://github.com/lfrank14
date: 03-16-2019
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plotly)
library(lurrdatviz)
library(paletteer)
library(colorblindr)
```

```{r fig2a_data}

facat <- facat
facat_reps <- facat %>% 
  select(subnum, contains("_train_")) %>% 
  filter(subnum < 33) # missing data for most recent subjects

collabels <- colnames(facat_reps[-1])

# Tidy for plotting  
facat_tidy <- facat_reps %>% 
  gather(key = "tmp", value = "rep_strength", -1) %>% 
  separate(col = "tmp", into = c("roi", NA, "rep_type")) %>% 
  filter(rep_type != "itemrep" &
           roi != "hip")

# Summarize for plotting
facat_sum <- groupsum(facat_tidy, y = rep_strength, g1 = roi, g2 = rep_type)

```

```{r fig1_v1_stats, inclue=FALSE}
# Compare against 0, for reference / annotations
results <- list()
for (i in 1:(ncol(facat_reps) - 1)) {
  results[[i]] <- t.test(facat_reps[i + 1])
}

pvals <- map(results, function(x)x$p.value)
sig_reps <- tibble(condition = collabels,
                   pval = pvals)
```

```{r fig2a_final}

fig2a <- ggplot(facat_sum, aes(x = factor(roi, levels = c("ahip","phip","mofc")), 
                      y = mean, fill = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                 position = position_dodge(width = .9), width = .2, color = "gray30") +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("steelblue","firebrick3"),
                    name = NULL, labels = c("Item", "Category")) +
  labs(y = "Cohen's d", x = NULL, caption = "* p < .05, + p < .1",
       title = "Incidental Learning") +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  scale_y_continuous(limits = c(-.015,.035)) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        legend.justification = c("right","top"),
        axis.text.x = element_text(size = 14, color = "black"),
        plot.caption = element_text(hjust = 1.75)) +
  annotate(geom = "text", label = "*", x = .78, y = .03, size = 10) +
  annotate(geom = "text", label = "+", x = 2.78, y = .026, size = 7)

```

```{r fig2b_data}
dicat <- dicat %>% 
  select(subnum, contains("avg"))

dicat_tidy <- dicat %>% 
  gather(key = "tmp", value = "mbeta", -subnum) %>%
  separate(col = "tmp", into = c("mdl_fit","roi",NA))

dicat_sum <- groupsum(dicat_tidy, y = mbeta, g1 = mdl_fit, g2 = roi)
```

```{r dicat_stats, include=FALSE}
nvar <- length(dicat) - 1

test_betas <- list()
for (i in 1:nvar) {
  test_betas[[i]] <- t.test(dicat[i+1])
  test_betas[[i]]$data.name <- colnames(dicat)[i+1]
}
```

```{r fig2b_final}

fig2b <- ggplot(dicat_sum, aes(x = factor(roi, levels = c("ahip","phip","mofc")), 
                      y = mean, fill = mdl_fit)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se),
                position = position_dodge(width = .9), width = .2, color = "gray30") +
  scale_fill_manual(values = c("steelblue","firebrick3"),
                    name = NULL, labels = c("Exemplar", "Prototype")) +
  labs(y = "Beta Estimate", x = NULL,
       caption = "* p < .05, + p < .1",
       title = "Feedback-based Learning") +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        legend.justification = "top",
        axis.text.x = element_text(size = 14, color = "black"),
        plot.caption = element_text(hjust = 1.75)) +
  ylim(c(-.2, .2)) +
  annotate(geom = "text", label = "*", x = 1.78, y = -.16, size = 10) +
  annotate(geom = "text", label = "*", x = 3.23, y = .19, size = 10) +
  annotate(geom = "text", label = "+", x = 1.23, y = .16, size = 7)

```

```{r fig.width=10, fig.height=5}
cowplot::plot_grid(fig2a, fig2b, labels = NULL)
```


+ Data: FACAT & DICAT
+ Figure A: Bar plot showing mean strength of item and category representations across different brain regions. 
+ Figure B: Bar plot showing the average correlation between the model fits (exemplar v. prototype) and activity in each brain region. 
+ Message: To show that during an incidental category learning, item representations are present in anterior hippocampus and VMPFC, but during feedback-based learning, category representations are present in anterior hippocampus and VMPFC
+ Audience: advisor and lab mates, research colleagues
+ Notes: We only really care about the strength of each type of representation as it compares to 0.


### Versions of Figure 2A

Below is the first version of Figure 2A. I knew pretty clearly what I wanted to do with this figure so only minor edits were made. I added white lines around the bars to make them pop. 

```{r fig2a_v1}

ggplot(facat_sum, aes(x = factor(roi, levels = c("ahip","phip","mofc")), 
                      y = mean, fill = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 0, color = "#343434", alpha = .5) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                 position = position_dodge(width = .9), width = .2, color = "#343434") +
  theme_minimal(base_size = 16) +
  scale_fill_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Item", "Category")) +
  labs(y = "Cohen's d", x = NULL) +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  scale_y_continuous(limits = c(-.015,.035)) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        #legend.position = c(.99, .99), 
        legend.justification = c("right","top"),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(color = "black"),
        plot.caption = element_text(hjust = -1)) +
  annotate(geom = "text", label = "*", x = .78, y = .03, size = 10)

```

I also tried flipping the x and y axis so that the bars were presented horizontally, however, I thought it was difficult to read. 

```{r fig2a_v2}
ggplot(facat_sum, aes(x = factor(roi, levels = c("ahip","phip","mofc")), 
                      y = mean, fill = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 0, color = "#343434", alpha = .5) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                 position = position_dodge(width = .9), width = .2, color = "#343434") +
  theme_minimal(base_size = 16) +
  scale_fill_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Item", "Category")) +
  labs(y = "Cohen's d", x = NULL) +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  scale_y_continuous(limits = c(-.015,.035)) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(size = .3),
        panel.grid.minor.x = element_line(size = .3),
        #legend.position = c(.99, .99), 
        legend.justification = c("right","top"),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(color = "black"),
        plot.caption = element_text(hjust = -1)) +
  annotate(geom = "text", label = "*", x = .71, y = .03, size = 10) +
  coord_flip()
```

### Versions of Figure 2B

As I created this plot, I thought it was too similar to the figure created from the FACAT data. However, as I started to put together the narrative, I decided it would be nice to compare with representations during incidental learning.  

```{r fig2b_v1}

ggplot(dicat_sum, aes(x = roi, y = mean, fill = mdl_fit)) +
  geom_hline(yintercept = 0, color = "#343434", alpha = .9) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se),
                position = position_dodge(width = .9), 
                width = .2, color = "#343434") +
  scale_fill_OkabeIto(name = NULL,
                      labels = c("Exemplar Fits", "Prototype Fits")) +
  labs(y = "Parameter Estimates", x = NULL) +
  scale_x_discrete(labels = c("Ant. HIP", "VMPFC", "Post. HIP")) +
  theme_minimal(base_size = 16) +
  theme(legend.justification = "top")
  
```

