---
title: "Visualization #3"
description: |
  During incidental learning, item representations in brain track category generalization.
author:
  - name: Lea Frank
    url: https://github.com/lfrank14
date: 03-17-2019
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lurrdatviz)
library(paletteer)
```


```{r fig3_data}
facat_sub <- facat %>% 
  filter(subnum < 33) %>% 
  select(subnum, postpre_mSim_Relevant, postpre_mSim_Irrelevant,
         catACC, catACCnew, catACCold, 
         CHR, FA, mFirstName, mSurname, mFullName,
         contains("_train_")) %>% 
  gather(key = "tmp", value = "rep_str", 12:23) %>% 
  separate(col = "tmp", into = c("roi", NA, "rep_type")) %>%
  filter(roi != "hip" & rep_type != "itemrep") %>% 
  mutate(roi = factor(roi, levels = c("ahip","phip","mofc"),
                      labels = c("Ant. HIP", "Post. HIP", "VMPFC")))

```

```{r fig3_final, fig.height=4.2, fig.width=10.5}
facat_sub %>% 
  ggplot(aes(x = rep_str, y = catACCnew, 
             color = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_point(alpha = .7) +
  stat_smooth(geom = "line", se = FALSE, alpha = .3, method = "lm",
              size = 1) +
  geom_smooth(data = filter(facat_sub, roi == "VMPFC" & rep_type == "itemrep2"),
              se = FALSE, method = "lm", size = 1) +
  facet_wrap(~roi) +
  theme_minimal(base_size = 18) +
  labs(x = "Strength of Representation \n(Cohen's D)",
       y = "Categorization Accuracy \n(new faces)") +
  scale_color_manual(values = c("steelblue","firebrick3"),
                     name = NULL, labels = c("Item", "Category")) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 16),
        panel.grid = element_line(size = .3),
        axis.text = element_text(color = "black"),
        axis.title.x = element_text(margin = margin(15,0,0,0, unit = "pt")),
        axis.title.y = element_text(margin = margin(0,15,0,0, unit = "pt")),
        strip.text = element_text(size = 18))
```

+ Data: FACAT
+ Figure: scatter plot between strength of item/category representations in different brain regions and category generalization performance
+ Message: To show that during incidental learning, item representations in the VMPFC tracked categorization accuracy for new faces, providing neural evidence suggesting that participants were using an exemplar strategy to generalize.  
+ Audience: advisor, lab mates, research colleagues


### Versions of Figure 3

When creating this figure, I had two ideas in mind. The first version would be a scatter plot showing the relationship between the strength of memory representations in each brain region and categorization performance. The second version would show a distribution of beta estimates when I regress the strength of memory representations onto categorization accuracy. 

```{r fig3_v1, fig.height=4.2, fig.width=10.5}
facat_sub %>% 
  ggplot(aes(x = rep_str, y = catACCnew, color = rep_type)) +
  geom_point(alpha = .8) +
  stat_smooth(geom = "line", se = FALSE, alpha = .3, method = "lm",
              size = 1) +
  geom_smooth(data = filter(facat_sub, roi == "VMPFC" & rep_type == "itemrep2"),
              se = FALSE, method = "lm", size = 1) +
  facet_wrap(~roi) +
  theme_minimal(base_size = 16) +
  labs(x = "Strength of Representation \n(Cohen's D)",
       y = "Categorization Accuracy \n(new faces)") +
  scale_color_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Category Rep.", "Item Rep.")) +
  theme(legend.position = "bottom",
        axis.text = element_text(color = "black"))
```

```{r fig3_v2_data, include=FALSE}
cat_regress <- lm(data = facat, 
                  catACCnew ~ ahip_train_itemrep2 + ahip_train_catrep +
                    phip_train_itemrep2 + phip_train_catrep +
                    mofc_train_itemrep2 + mofc_train_catrep)

summary(cat_regress)

mdl <- broom::tidy(cat_regress) %>% 
  mutate(ci_lower = estimate - (1.96)*std.error,
         ci_upper = estimate + (1.96)*std.error) %>% 
  filter(term != "(Intercept)")

mdl$term <- factor(mdl$term, levels = c("ahip_train_itemrep2", "ahip_train_catrep",
                                        "phip_train_itemrep2", "phip_train_catrep",
                                        "mofc_train_itemrep2", "mofc_train_catrep"))  
```

```{r fig3_v2}

term_color <- ifelse(mdl$p.value < .05, "#15C18C", "#6E6E6E") 

ggplot(mdl, aes(x = term, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = .2,
                alpha = .9, color = "#6E6E6E") +
  geom_point(alpha = .8, color = "#6E6E6E") +
  geom_errorbar(data = filter(mdl, term == "mofc_train_itemrep2"),
                aes(ymin = ci_lower, ymax = ci_upper), 
                width = .2, color = "#15C18C") +
  geom_point(data = filter(mdl, term == "mofc_train_itemrep2"), 
             color = "#15C18C") +
  coord_flip() +
  theme_minimal(base_size = 14) +
  labs(x = NULL, y = "Beta Estimate",
       caption = "Figure 3. Item representations in the VMPFC predict category generalization") +
  scale_x_discrete(limits = rev(levels(mdl$term)),
                   labels = c("VMPFC Cat Rep.", "VMPFC Item Rep.",
                              "Post. HIP Cat Rep.", "Post. HIP Item Rep.",
                              "Ant. HIP Cat Rep.", "Ant. HIP Item Rep.")) +
  theme(axis.text.y = element_text(color = rev(term_color)),
        plot.caption = element_text(hjust = 1.5, size = 14))

```

I had a hard time deciding which version of figure 3 that I wanted to go with for the final product. Ultimately, I chose to stick with version 1 so that the viewers could get an idea of the distribution of the data. I felt this was particularly important as the data is exploratory. Moreover, I thought version 1 would fit well with the formatting of the other two figures and maintain a coherent narrative. 